State of the Practice for Medical Imaging Software Based on Open Source Repositories

### 2024-10-29 StateOfPractice/Papers/SOP_MI_CritReviewsBioMedEng/SOP_MI_OSR.tex
Updates to Section 7

Our observations of an informally defined process, with elements of agile
methods, matches what has been observed for research software in general.
Scientific developers naturally use an agile philosophy \cite{AckroydEtAl2008,
CarverEtAl2007, EasterbrookAndJohns2009, Segal2005, HeatonAndCarver2015}, or an
amethododical process \cite{Kelly2013}, or a knowledge acquisition driven
process \cite{Kelly2015}.  A waterfall-like process can work for research
software \cite{Smith2016}, especially if the developers work iteratively and
incrementally, but externally document their work as if they followed a
rationale design process \cite{parnas1986rational}.

No interviewee introduced any strictly defined project management process. The
most common approach was following the issues, such as bugs and feature
requests. Additionally, the \textit{3D Slicer} team had weekly meetings to
discuss the goals for the project; the \textit{INVESALIUS 3} team relied on the
GitHub process for their project management; the \textit{ITK-SNAP} team had a
fixed six-month release pace; only the interviewee from the \textit{OHIF} team
mentioned that the team has a project manager; the \textit{3D Slicer} team and
\textit{BioImage Suite Web} team do nightly builds and tests. The \textit{OHIF}
developer believes that a better project management process can improve junior
developer efficiency while also improving internal and external communication.

We identified the use of unit testing in less than half of the 29 projects. On
the other hand, the interviewees believed that testing (including usability
tests with users) was the top solution to improve correctness, usability, and
reproducibility.  This level of testing matches what was observed for LBM
software \cite{Michalski2021} and is apparently greater than the level of
testing for ocean modelling software \cite{JungEtAl2022}.

As the observed artifacts in Table~\ref{artifactspresent} show, none of the 29
projects emphasize documentation. None of them had theory manuals, although we
did identify a road map in the \textit{3D Slicer} project.  We did not find
requirements specifications. Table~\ref{tab_opinion_doc} summarizes
interviewees' opinions on documentation. Interviewees from each of the eight
projects thought that documentation was essential to their projects, and most of
them said that it could save their time to answer questions from users and
developers. Most of them saw the need to improve their documentation, and only
three of them thought that their documentations conveyed information clearly
enough. Nearly half of developers also believed that the lack of time prevented
them from improving documentation.

\begin{table}[!ht]
\centering
\begin{tabular}{ll}
\toprule
Opinion on Documentation & Num.\ Ans. \\ 
\midrule
Documentation is vital to the project & 8 \\
Documentation of the project needs improvements & 7 \\
Referring to documentation saves time to answer questions & 6 \\
Lack of time to maintain good documentation & 4 \\
Documentation of the project conveys information clearly & 3 \\
Coding is more fun than documentation & 2 \\
Users help each other by referring to documentation & 1 \\ 
\bottomrule
\end{tabular}
\caption{Opinions on documentation by the numbers of interviewees with the
answers}
\label{tab_opinion_doc}
\end{table}

>>>>

Our observations of an informally defined process, with elements of agile
methods, matches observations for research software in general.
Scientific developers naturally use an agile philosophy \cite{AckroydEtAl2008,
CarverEtAl2007, EasterbrookAndJohns2009, Segal2005, HeatonAndCarver2015}, or an
amethododical process \cite{Kelly2013}, or a knowledge acquisition driven
process \cite{Kelly2015}.  A waterfall-like process can work for research
software \cite{Smith2016}, especially if the developers work iteratively and
incrementally, but externally document their work as if they followed a
rationale design process \cite{parnas1986rational}.

No interviewee introduced any strictly defined project management process. The
most common approach was following the issues, such as bugs and feature
requests. The \textit{3D Slicer} team had weekly meetings to
discuss the goals for the project. The \textit{INVESALIUS 3} team relied on the
GitHub process for their project management while the \textit{ITK-SNAP} team had a
fixed six-month release pace. Only the interviewee from \textit{OHIF} 
mentioned the team having a project manager. The \textit{3D Slicer} team and
\textit{BioImage Suite Web} team do nightly builds and tests. The \textit{OHIF}
developer believes that a better project management process can improve junior
developer efficiency while also improving internal and external communication.

We identified the use of unit testing in less than half of the 29 projects even
though the interviewees believed that testing (including usability tests with
users) was the top approach to improve correctness, usability, and
reproducibility.  This level of testing matches what was observed for LBM
software [citation redacted for blind review]
%\cite{Michalski2021} 
and is apparently greater than the level of
testing for ocean modelling software \cite{JungEtAl2022}.

Interviewees from all of the eight projects thought that documentation was
essential to their projects, and most of them (7 of 8) said that it could save
their time to answer questions from users and developers. Half of them (4 of 8)
saw the need to improve their documentation, and only three of them thought that
their documentations conveyed information clearly enough.
